Is high social class always beneficial for survival? A study of northern Sweden 1801--2013 \footnote{Short; "Is high social class beneficial for survival?"}

### 2018-10-10 future.Rmd
remove some errors

```{r readalldata, eval = FALSE}
source("R/get.file.R")
##skume <- get.file("skume")
load("data/skume.rda")
skume <- skume[order(skume$id, skume$enter), ]
skume$id <- paste("pop", skume$id, sep = "_")
##
skume1 <- age.window(skume, c(40, 65))  ###### OJOJOJ!!!!!!!!!!!!!!!!!!!!!!!!!!! (40, 65)
skume1 <- oneSpell(skume1)
##
skume2 <- age.window(skume, c(65, 100))
skume2 <- oneSpell(skume2)
##################################################
##foli <- get.file("foli")
load("data/foli.rda")
foli$civst[is.na(foli$civst)] <- "unmarried"
foli <- foli[order(foli$id, foli$enter), ]
foli$AR <- foli$ULORSAK <- NULL
##
foli1 <- age.window(foli, c(40, 65))
foli1 <- oneSpell(foli1)
##
foli2 <- age.window(foli, c(65, 100))
foli2 <- oneSpell(foli2)
##################################################
source("CreData/R/addPer.R")
skume1 <- addPer(skume1, cuts = seq(1801, 1951, by = 25))
skume2 <- addPer(skume2, cuts = seq(1801, 1951, by = 25))
##
foli1 <- addPer(foli1, cuts = c(1976, 1986, 1996, 2005, 2014))
foli2 <- addPer(foli2, cuts = c(1976, 1986, 1996, 2005, 2014))
##################################################
vb1 <- rbind(skume1, foli1)
vb1$period <- factor(vb1$period)
vb1Tab <- toTpch(vb1[, !(colnames(vb1) %in% c("id", "birthdate"))], cuts = seq(40, 65, by = 1))
##
vb2 <- rbind(skume2, foli2)
vb2$period <- factor(vb2$period)
vb2Tab <- toTpch(vb2[, !(colnames(vb2) %in% c("id", "birthdate"))], cuts = seq(65, 100, by = 1))
vb <- rbind(vb1, vb2)
vb$exposure <- vb$exit - vb$enter # For statistics and figures purposes
vbTab <- rbind(vb1Tab, vb2Tab) # NEW
```

@@

```{r hazcalw}
source("R/getWeights.R")
labb <- levels(vbTab$period)
nper <- length(labb)
his <- levels(vb1Tab$hisclass)
nhis <- length(his)
##
resm <- matrix(0, nrow = nhis, ncol = nper)
pdm <-  matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb1Tab[vb1Tab$sex == "male" & vb1Tab$period == labb[i], ]
    ##dat <- vb1[vb1$sex == "male" & vb1$period == labb[i], ]
    ##dat$wght <- getWeights(dat)
    ##fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    ##x <- survfit(fit)
    for (j in seq_along(his)){
        fit <- tpchreg(event ~ offset(log(exposure)) + age, data = dat[dat$hisclass == his[j], ], cuts = seq(40, 65, by = 1))
        haz <- c(0, fit$hazards, fit$hazards[length(fit$hazards)])
        pdm[j, i] <- ppch(65, cuts = fit$cuts, levels = haz)
    }
    ##resm[, i] <- summary(x, rmean = 65)$table[, "*rmean"]
    ##pdm[, i] <- 1 - x$surv[cumsum(x$strata)]
}
##
resw <- matrix(0, nrow = nhis, ncol = nper)
pdw <-  matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb1Tab[vb1Tab$sex == "female" & vb1Tab$period == labb[i], ]
    ##dat <- vb1[vb1$sex == "female" & vb1$period == labb[i], ]
    ##dat$wght <- getWeights(dat)
    ##fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    ##x <- survfit(fit)
    for (j in seq_along(his)){
        fit <- tpchreg(event ~ offset(log(exposure)) + age, data = dat[dat$hisclass == his[j], ], cuts = seq(40, 65, by = 1))
        haz <- c(0, fit$hazards, fit$hazards[length(fit$hazards)])
        pdw[j, i] <- ppch(65, cuts = fit$cuts, levels = haz)
    }
    ##resw[, i] <- summary(x, rmean = 65)$table[, "*rmean"]
    ##pdw[, i] <- 1 - x$surv[cumsum(x$strata)]
}

@@

# vb3064$social_class <- vb3064$hisclass
# vb3064$marital_status <- vb3064$civst
# vb3064$urban_or_rural <- vb3064$urban
# vb3064$age_group <- vb3064$age
fitt <- tpchreg(event ~ offset(log(exposure)) + age + urban + civst + sex * hisclass * period, cuts = 40:65,
             data = vb1Tab)
dr <- drop1(fitt, test = "Chisq")
# ##dr
```

@@

fittold <- coxph(Surv(enter, exit, event) ~ hisclass * period * sex + civst + urban, data = vb2)
drold <- drop1(fittold, test = "Chisq")
```

@@

```{r hazcalw2}
nper <- length(labb)
his <- levels(vb1$hisclass)
nhis <- length(his)
##
resm <- matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb2[vb2$sex == "male" & vb2$period == labb[i], ]
    dat$wght <- getWeights(dat)
    fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    resm[, i] <- summary(survfit(fit), rmean = 100)$table[, "*rmean"]
}
##
resw <- matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb2[vb2$sex == "female" & vb2$period == labb[i], ]
    dat$wght <- getWeights(dat)
    fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    resw[, i] <- summary(survfit(fit), rmean = 100)$table[, "*rmean"]
}

>>>>

```{r readalldata, eval = FALSE}
source("R/get.file.R")
##skume <- get.file("skume")
load("data/skume.rda")
skume <- skume[order(skume$id, skume$enter), ]
skume$id <- paste("pop", skume$id, sep = "_")
##
skume1 <- age.window(skume, c(40, 65))  ###### OJOJOJ!!!!!!!!!!!!!!!!!!!!!!!!!!! (40, 65)
skume1 <- oneSpell(skume1)
##
skume2 <- age.window(skume, c(65, 100))
skume2 <- oneSpell(skume2)
##################################################
##foli <- get.file("foli")
load("data/foli.rda")
foli$civst[is.na(foli$civst)] <- "unmarried"
foli <- foli[order(foli$id, foli$enter), ]
foli$AR <- foli$ULORSAK <- NULL
##
foli1 <- age.window(foli, c(40, 65))
foli1 <- oneSpell(foli1)
##
foli2 <- age.window(foli, c(65, 100))
foli2 <- oneSpell(foli2)
##################################################
source("CreData/R/addPer.R")
skume1 <- addPer(skume1, cuts = seq(1801, 1951, by = 25))
skume2 <- addPer(skume2, cuts = seq(1801, 1951, by = 25))
##
foli1 <- addPer(foli1, cuts = c(1976, 1986, 1996, 2005, 2014))
foli2 <- addPer(foli2, cuts = c(1976, 1986, 1996, 2005, 2014))
##################################################
vb1 <- rbind(skume1, foli1)
vb1$period <- factor(vb1$period)
vb1Tab <- toTpch(vb1[, !(colnames(vb1) %in% c("id", "birthdate"))], cuts = seq(40, 65, by = 5))
##
vb2 <- rbind(skume2, foli2)
vb2$period <- factor(vb2$period)
vb2Tab <- toTpch(vb2[, !(colnames(vb2) %in% c("id", "birthdate"))], cuts = seq(65, 100, by = 5))
vb <- rbind(vb1, vb2)
vb$exposure <- vb$exit - vb$enter # For statistics and figures purposes
vbTab <- rbind(vb1Tab, vb2Tab) # NEW
```

@@

```{r hazcalw}
source("R/getWeights.R")
labb <- levels(vbTab$period)
nper <- length(labb)
his <- levels(vb1Tab$hisclass)
nhis <- length(his)
##
resm <- matrix(0, nrow = nhis, ncol = nper)
pdm <-  matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb1Tab[vb1Tab$sex == "male" & vb1Tab$period == labb[i], ]
    ##dat <- vb1[vb1$sex == "male" & vb1$period == labb[i], ]
    ##dat$wght <- getWeights(dat)
    ##fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    ##x <- survfit(fit)
    for (j in seq_along(his)){
        fit <- tpchreg(event ~ offset(log(exposure)) + age, data = dat[dat$hisclass == his[j], ], cuts = seq(40, 65, by = 5))
        haz <- c(0, fit$hazards, fit$hazards[length(fit$hazards)])
        pdm[j, i] <- ppch(65, cuts = fit$cuts, levels = haz)
    }
    ##resm[, i] <- summary(x, rmean = 65)$table[, "*rmean"]
    ##pdm[, i] <- 1 - x$surv[cumsum(x$strata)]
}
##
resw <- matrix(0, nrow = nhis, ncol = nper)
pdw <-  matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb1Tab[vb1Tab$sex == "female" & vb1Tab$period == labb[i], ]
    ##dat <- vb1[vb1$sex == "female" & vb1$period == labb[i], ]
    ##dat$wght <- getWeights(dat)
    ##fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    ##x <- survfit(fit)
    for (j in seq_along(his)){
        fit <- tpchreg(event ~ offset(log(exposure)) + age, data = dat[dat$hisclass == his[j], ], cuts = seq(40, 65, by = 5))
        haz <- c(0, fit$hazards, fit$hazards[length(fit$hazards)])
        pdw[j, i] <- ppch(65, cuts = fit$cuts, levels = haz)
    }
    ##resw[, i] <- summary(x, rmean = 65)$table[, "*rmean"]
    ##pdw[, i] <- 1 - x$surv[cumsum(x$strata)]
}

@@

# vb3064$social_class <- vb3064$hisclass
# vb3064$marital_status <- vb3064$civst
# vb3064$urban_or_rural <- vb3064$urban
# vb3064$age_group <- vb3064$age
fitt <- tpchreg(event ~ offset(log(exposure)) + age + urban + civst + sex * hisclass * period, cuts = seq(40, 65, by = 5),
             data = vb1Tab)
dr <- drop1(fitt, test = "Chisq")
# ##dr
```

@@

##fittold <- coxph(Surv(enter, exit, event) ~ hisclass * period * sex ##+ civst + urban, data = vb2)
fittold <- tpchreg(event ~ offset(log(exposure)) + age + hisclass * period * sex + civst + urban, cuts = seq(65, 100, by = 5), data = vb2Tab)
drold <- drop1(fittold, test = "Chisq")
```

@@

```{r hazcalw2}
nper <- length(labb)
his <- levels(vb2Tab$hisclass)
nhis <- length(his)
##
resm <- matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb2Tab[vb2Tab$sex == "male" & vb2Tab$period == labb[i], ]
    dat$wght <- getWeights(dat)
    for (j in seq_along(his)){
        fit <- tpchreg(event ~ offset(log(exposure)) + age, 
                       cuts = seq(65, 100, by = 5), 
                       data = dat[dat$hisclass == his[i], ])
        haz <- c(0, fit$hazards, fit$hazards[length(fit$hazards)])
        resm[j, i] <- mpch(levels = haz, cuts = 65:100)
    }
    ##fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    ##resm[, i] <- summary(survfit(fit), rmean = 100)$table[, "*rmean"]
}
##
resw <- matrix(0, nrow = nhis, ncol = nper)
for (i in 1:nper){
    dat <- vb2Tab[vb2Tab$sex == "female" & vb2Tab$period == labb[i], ]
    dat$wght <- getWeights(dat)
        for (j in seq_along(his)){
        fit <- tpchreg(event ~ offset(log(exposure)) + age, 
                       cuts = seq(65, 100, by = 5), 
                       data = dat[dat$hisclass == his[i], ])
        haz <- c(0, fit$hazards, fit$hazards[length(fit$hazards)])
        resw[j, i] <- mpch(levels = haz, cuts = 65:100)
    }

    ##fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = dat)
    ##resw[, i] <- summary(survfit(fit), rmean = 100)$table[, "*rmean"]
}
